### 2015-01-28 ASP.tex
improvement of ASP + update

\emph{Answer Set Programming} (ASP) is a declarative programming paradigm with semantics known as the \emph{semantics of answer sets}.
This paradigm allows the programmer to specify what the problem to be solved is, and not how to solve it.
ASP programs are written in AnsProlog* (short for "Answer Set Programming in Logic" with superscript:~*).
These programs are composed of a set of \emph{facts} and a set of \emph{rules}, as defined below, from which other facts can be derived.
A consistent set of facts that can be derived from a program using the rules is called \emph{answer set} for the program.
The sets of possible responses to an AnsProlog* program are calculated with a program called a solver.

\subsubsection{Logic program}
An ASP program is composed of rules, each one being an ordered pair~\cite{baral2003knowledge}:
\begin{equation} \label{eq1ASP}
head \leftarrow body.
\end{equation}
with $head$ and $body$ two sets of literals.

This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving.

\subsubsection{The syntax and semantic of ASP}
\label{sectionSyntaxeASP}
\begin{enumerate}
\item \textbf{The alphabet}

\noindent{ According to \cite{baral2003knowledge}, the alphabet of the axioms (or simply called the \emph{alphabet}) of ASP is composed of seven classes of symbols:
\begin{itemize}
\item Variables,
\item Constants,
\item Symbols of functions,
\item Symbols of predicates,
\item Connectors,	
\item Symbols of punctuation,
\item The special symbol $\bot$
\end{itemize}
}
All these classes vary from one alphabet to another, but the sets of $ 5^{th} $ and $ 6^{th} $ classes (Connectors and symbols of punctuation) are defined as follows:
\begin{itemize}
  \item Connectors with \{$\neg$, or, $\leftarrow$, not, ‘,’\}
  \item Symbols of punctuation with \{‘(’, ‘)’, ‘.’\}
\end{itemize}

\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
\repEB{à revoir}
The other classes remain constant as they meet some informal agreement. In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).

\annotMF{Je ne suis pas sûr que l'explication sur les prédicats soit pertinente. Si on la garde, il faudrait la reformuler.}
What seems a little fuzzy is the concept of predicate. Indeed the word predicate can be an innovation of the new grammar. Let's consider the following sentence: 
\begin{tabbing}
 \textit{Socrates was an athenian.} 
\end{tabbing}
\textit{Socrate} is the subject and \textit{was an athenian} is the predicat. This sentence can be noted in ASP by :
\begin{tabbing}
 \texttt{athenian(Socrate) $\leftarrow ~ \top$. }\hspace{2cm} with \texttt{$\top$}  is the symbol of True.
\end{tabbing}

\item \textbf{The rules}
As previously mentioned, a rule has the following form:
\texttt{
\begin{equation*} \label{eq3ASP}
 A ~ \leftarrow ~ B^+, ~ \ASPnot B^-.
\end{equation*}
}

where the part on the left of the arrow is called $ head$
and the right part is called $body$.
This rule is then equivalent to:
\texttt{
\begin{equation*} \label{eq2ASP}
l_{0} \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
\end{equation*}
}
And this rule states:
\[
\left.
    \begin{array}{ll}
        \text{If} ~ \texttt{$l_{1},...,l_{m}$}  ~ \text{are \textbf{true}}\\
        \text{and} ~ \texttt{$l_{m+1},...,l_{n}$} ~ \text{are \textbf{false}}
    \end{array}
\right \} \text{Then} ~ \texttt{$ l_{0} $} ~ \text{is } \textbf{true} 
\]
with \texttt{$l_{i}$} literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
Therefore, $head$ is also called the \emph{conclusion} of the rule,
and \texttt{body} is its \emph{premise}.

There are some special cases of rules \cite{Vladimir,baral2003knowledge}:
\begin{itemize}
\item A rule is said \textbf{constant} or \emph{ground} if all of its literals are constants;
\item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
  It can be written without the central arrow ($\leftarrow$) or with a body equal to $\top$:
  \texttt{
\begin{equation} 
l_{0} \leftarrow \top. \qquad \text{or} \qquad l_{0}.
 \label{eq5ASP}
\end{equation}
}

@@

\end{enumerate}

\subsubsection{Modeling of a problem with ASP}

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}:
\begin{itemize}
\item it is a simplification of a given system,
\item it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties... 
\end{itemize}
This concept model can address another angle issues related to representation process.

\begin{enumerate}
\item\textbf{Modeling steps}
The modeling process can be regarded as a special form of representation whose operations are detailed in \cite{baral2003knowledge}.
Logical ASP programs follow the strategy of “generate and test”.
This strategy includes four steps:
\begin{itemize}
\item Enumerate with facts;
\item Explain with rules;
\item Generate all the possibilities with cardinalities; and finally,
\item Filter with constraints.
\end{itemize}

\item \textbf{Solving an ASP program}
The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
that usually work together:
first the grounder is used to remove variables in order to achieve an equivalent but constant program,
then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
Amongst all available grounders for ASP, we can name
\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
and the solvers include
\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...

\reformuler{
The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations. Both systems employ grounding algorithms that work quickly and simplify the program.
}
For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

\end{enumerate}

>>>>

\emph{Answer Set Programming} (ASP) is a declarative programming paradigm with semantics known as the \emph{semantics of answer sets}.
This paradigm allows the programmer to specify what the problem to be solved is, and not how to solve it.
ASP programs are written in AnsProlog* \cite{sureshkumar2006ansprolog} (short for "Answer Set Programming in Logic" with superscript:~*).
These programs are composed of a set of \emph{facts} and a set of \emph{rules}, as defined below, from which other facts can be derived.
A consistent set of facts that can be derived from a program using the rules is called \emph{answer set} for the program.
The sets of possible responses to an AnsProlog* program are calculated with a program called a solver.

\subsubsection{The syntax and semantic of ASP}
\label{sectionSyntaxeASP}
\begin{enumerate}

\item \textbf{The rules:}
As previously mentioned, a rule has the following form:
\texttt{
\begin{equation} \label{eq1ASP}
 head \leftarrow body.
\end{equation}
}
This rule is equivalent to:
\begin{equation} \label{eq2ASP}
l_{0} \leftarrow l_{1},...,l_{m}, \ASPnot l_{m+1},..., \ASPnot l_{n}.
\end{equation}
And this rule states:
\[
\left.
    \begin{array}{ll}
        \text{If} ~ l_{1},...,l_{m}  ~ \text{are \textbf{true}}\\
        \text{and} ~ l_{m+1},...,l_{n} ~ \text{are \textbf{false}}
    \end{array}
\right \} \text{Then} ~ l_{0} ~ \text{is } \textbf{true} 
\]
with $l_{i}$ literals for all $i \in \llbracket 1 ; n \rrbracket$, and $0 \leq m \leq n$.
Therefore, "$head$" is also called the \emph{conclusion} of the rule,
and "$body$" is its \emph{premise}.

There are some special cases of rules \cite{Vladimir,baral2003knowledge}:
\begin{itemize}
\item A  \textbf{ground} rule, is a rule  where all its literals are constants;
\item A \textbf{fact} or a \textbf{reality} is a rule with an empty body.
  It can be written without the central arrow ($\leftarrow$) or with a $body$ equals to $\top$, and $\top$  is the symbol of True.
\begin{equation} 
l_{0} \leftarrow \top. \qquad \text{or} \qquad l_{0}.
 \label{eq5ASP}
\end{equation}

@@

\item \textbf{The alphabet}

According to \cite{baral2003knowledge}, the alphabet of ASP is composed of seven classes of symbols: Variables, start with a capital letter. Constants, symbols of functions, symbols of predicates, these 3 classes follow the same rule but they begin with lowercase letter. Connectors: $\neg$, or, $\leftarrow$, not, ‘,’. Symbols of punctuation :(‘(’, ‘)’, ‘.’) and finally the special symbols : $\bot$ and $\top$.

The word predicate can be an innovation of the new grammar. Let's consider the following sentence to more explain it: 
\begin{tabbing}
 \textit{tweety is a bird.} 
\end{tabbing}
"\textit{tweety}" is a constant and "\textit{is a bird}" is a predicat. This sentence can be noted in ASP by :
\begin{tabbing}
 bird(tweety) $\leftarrow ~ \top$. \hspace{1cm} or \hspace{2cm} $bird(tweety) .$
\end{tabbing} 
\end{enumerate}

\subsubsection{Modeling and solving of a problem with ASP: }

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}: In fact it is a simplification of a given system and it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties... This concept model can address another angle issues related to representation process. Logical ASP programs follow the strategy of “generate and test”. Indeed, the modeling process can be regarded as a special form of representation. First, the system should be presented  with facts. Second it have to be explain with rules. Third, it should be possible to generate all possibilities using cardinalities. Finally, a filter will be done thanking to the constraints and only the sets of predicates satisfying all these conditions will be displayed which are considered "answer sets".

We denote that the solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
that usually work together:
first the grounder is used to remove variables in order to achieve an equivalent but constant program,
then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
Amongst all available grounders for ASP, we can name
\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
and the solvers include
\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...

For the work presented in this paper, we worked with \texttt{Clingo} (verion 3) which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-28 PH.tex
improvement of ASP + update

\modMF{%
The PH is therefore a restriction of synchronous automata, where each transition
changes the local state of exactly one automaton,
and is triggered by the local states of at most two distinct automata.
This restriction in the form of the actions was chosen to permit
the development of efficient static analysis methods
based on abstract interpretation~\cite{PMR12-MSCS}.
}%

>>>>

The PH is therefore a restriction of synchronous automata, where each transition
changes the local state of exactly one automaton,
and is triggered by the local states of at most two distinct automata.
This restriction in the form of the actions was chosen to permit
the development of efficient static analysis methods
based on abstract interpretation~\cite{PMR12-MSCS}.

### 2015-01-28 comparaison.tex
improvement of ASP + update

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al. \cite{roccaasp} which was developed also on ASP.
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
\annotMF{Je propose plutôt : “ASP for CTL model-checking” qui pourrait se simplifier en “ASP-CTL” dans le tableau.}
Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transition network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

@@

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & $ \lbrack 5mn ; \infty \rbrack$ & 4m27.93s \\
  \hline
   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

@@

\item \textbf{\textsc{GINsim}} is a software for the edition, simulation and analysis
of gene interaction networks.
It allows to compute all reachable stable states instantly;
however, it is not possible to compute all stable states independently from the initial state.
Regarding the reachability problem, \textsc{GINSim} only allows to check
full states, because of its approach which consists in computing
all the state-transition graph
\todo{Seulement la partie accessible ou l'ensemble du graphe ?} \repEB{Je pense que oui mais je n'arrive pas à le m'assurer car il m'a fallut les états initiaux pour tester l'atteignabilité à ces états stables}
and then check if there exists a path between two given states.
Therefore, it was not possible to perform reachability checks on partial states.
Small state-transition graphs can also be displayed by this tool.

@@

\item \textbf{PINT}
is a library gathering tools and translators related to the PH.
It should be noted that Pint contains the only reachability analysis
developed so far natively for the Process Hitting,
before the method proposed in this paper.
It consists in an approximation that avoids to compute the state-transition graph;
It is thus ensured to be really efficient, which explains the fastest results,
but at the cost of possibly terminating without being conclusive (although this is rare).
However, it is not designed for goals consisting of a full state, which
may trigger an exponential process to reach an answer,
which explains the high computing times for some of them.
\todo{Ref vers le tableau}
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
\end{itemize}

>>>>

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method for CTL model-checking of Rocca's et al. \cite{roccaasp} which was developed also on ASP but for transition states networks.
Each method use a specific kind of representation\footnote{The transalation from PH to other models is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transitions network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

@@

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & \textsc{ASP-CTL} & \textsc{Pint} & \textsc{libddd} & \textsc{GINsim} & \textsc{ASP-PH} \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & out & 4m27.93s \\
  \hline
   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method denoted by \textsc{ASP-CTL}, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method \textsc{ASP-PH} developped in ASP for Process Hitting networks.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

@@

\item \textbf{\textsc{GINsim}} is a software for the edition, simulation and analysis
of gene interaction networks.
It allows to compute all reachable stable states instantly;
however, it is not possible to compute all stable states independently from the initial state.
Regarding the reachability problem, \textsc{GINSim} only allows to check
full states, because of its approach which consists in computing
all the state-transition graph and then check if there exists a path between two given states.
Therefore, it was not possible to perform reachability checks on partial states.
Small state-transition graphs can also be displayed by this tool.

@@

\item \textbf{PINT}
is a library gathering tools and translators related to the PH.
It should be noted that Pint contains the only reachability analysis
developed so far natively for the Process Hitting,
before the method proposed in this paper.
It consists in an approximation that avoids to compute the state-transition graph;
It is thus ensured to be really efficient, which explains the fastest results,
but at the cost of possibly terminating without being conclusive (although this is rare).
However, it is not designed for goals consisting of a full state, which
may trigger an exponential process to reach an answer,
which explains the high computing times for some of them \ref{tab:reachability}
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
\end{itemize}

### 2015-01-28 intro.tex
improvement of ASP + update

\todo{à revoir} \\
As regulatory phenomena play a crucial role in biological systems, they need to
be studied accurately. Biological Regulatory Networks (BRNs) consist in sets
of either positive or negative mutual effects between the components. With the
purpose of analyzing these systems, they are often modeled as graphs which make
it possible to determine the possible evolutions of all the interacting components of the system. Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels. A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system. This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using approached methods computing approximations of the dynamics~\cite{PMR12-MSCS}. Furthermore, PH is suitable, according to the precision of the available information, to model BRNs with different levels of abstraction by capturing the most general dynamics.

The objectives of the work presented in this paper are the following.
Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fix points which is a specific attractor with only one state \cite{wuensche1998genomic}).
For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fix points.
The second phase of our work consists in computing the dynamics. It consists in determining from a known initial state the possible next states of the PH model. Then we verify if we can reach a specific state for one or several component (genes or proteins). The results are ensured to respect strictly the PH dynamics.

Our contribution is from the results that allowed to determine the stable states, we propose to evaluate the benefits of the Answer Set Programming (ASP) \cite{baral2003knowledge} to compute them. ASP has been proven efficient to tackle models with a large number of components and parameters. Our aim here is to assess its potential w.r.t. the computation of some dynamical properties of the PH model. In this paper, we show that ASP turns out to be effective for these enumerative searches which justifies its use. The benefit of our approach is that it makes possible to get the minimal paths to reach our goal(s) also we can verify if it is possible after a given number of steps. %Furthermore, we claim that we are able to deal with large biological networks (42 sorts).

\annotMF{Je pense que dans toute l'intro, il faut remettre ASP au centre car c'est bien lui qui rend notre contribution particulière. Il ne faudrait pas séparer Objectifs/Contributions car les deux se recoupent, ou trouver un moyen de parler d'ASP avant de présenter le travail sur les points fixes et la dynamique.}

Section \ref{defs} defines the paradigm of Answer Set Programming,
and the Process Hitting framework.
We then use ASP in section \ref{fixpoint} to tackle the enumeration
of all fix points in a model,
and in section \ref{dynamics} in order to compute the dynamics of a model
and check reachability properties.
Finally, a comparison of the methods developed in this paper is
proposed in section \ref{comparison},
and in section \ref{ccl} we conclude and talk about future work.

>>>>

As regulatory phenomena play a crucial role in biological systems, they need to
be studied accurately. Biological Regulatory Networks (BRNs) consist in sets
of either positive or negative mutual effects between the components. With the
purpose of analyzing these systems, they are often modeled as graphs which make
it possible to determine the possible evolutions of all the interacting components of the system. Indeed, in order to address the formal checking of dynamical properties within very large BRNs, we recently introduced a new formalism, named the “Process Hitting” (PH) \cite{PMR10-TCSB}, to model concurrent systems having components with a few qualitative levels. A PH describes, in an atomic manner, the possible evolutions of a “process” (representing one component at one level) triggered by the hit of at most one other “process” in the system. This particular structure makes the formal analysis of BRNs with hundreds of components tractable by using approached methods computing approximations of the dynamics~\cite{PMR12-MSCS}. Furthermore, PH is suitable, according to the precision of the available information, to model BRNs with different levels of abstraction by capturing the most general dynamics.

The objectives of the work presented in this paper are the following.
Firstly, we show that starting from one PH model, it is possible to find all possible stable states (fixed points which is a specific attractor with only one state \cite{wuensche1998genomic}).
For this, we perform an exhaustive search of the possible states, that is, combinations of processes with one process from each sort, and then check which ones are fixed points.
The second phase of our work consists in computing the dynamics. It consists in determining from a known initial state the possible next states of the PH model. Then we verify if we can reach a specific state for one or several component (genes or proteins). The results are ensured to respect strictly the PH dynamics.

Our contribution is from the results that allowed to determine the stable states, we propose to evaluate the benefits of the Answer Set Programming (ASP) \cite{baral2003knowledge} to compute them. ASP has been proven efficient to tackle models with a large number of components and parameters. Our aim here is to assess its potential w.r.t. the computation of some dynamical properties of the PH model. In this paper, we show that ASP turns out to be effective for these enumerative searches which justifies its use. The benefit of our approach is that it makes possible to get the minimal paths to reach our goal(s) also we can verify if it is possible after a given number of steps.

Section \ref{defs} defines the paradigm of Answer Set Programming,
and the Process Hitting framework.
We then use ASP in section \ref{fixpoint} to tackle the enumeration
of all fixed points in a model,
and in section \ref{dynamics} in order to compute the dynamics of a model
and check reachability properties.
Finally, a comparison of the methods developed in this paper is
proposed in section \ref{comparison},
and in section \ref{ccl} we conclude and talk about future work.

### 2015-01-28 reachability.tex
improvement of ASP + update

\subsection{Reachability verification}
In this section, we focus on the reachability of set of processes which corresponds to the reachability question (see definition \ref{def:reachability}):
“Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
\annotMF{En fin de rédaction, il faudrait voir à peut-être décaler les définitions de la section “Dynamic properties” afin de faciliter la lecture.}
\repEB{Ok}
We now want to adapt the code of the dynamics computation of previous section in order to resolve this reachability problem.
For this, we first use a predicate to list the processes we want to reach, called \texttt{goal}, and we add as many rules of the following form as there are objective processes:
\begin{lstlisting}
goal(activeProcess("a",1)). %\label{c1}
\end{lstlisting}
Then, the literal \texttt{satisfiable(F, T)} 
checks if a given process \texttt{F} of the goal
is contained in the state of step \texttt{T},
as defined in the rule of line \ref{c2}.
Else the answer will be eliminated by the constraint \ref{c2-a} which verifies if all processes of the goal are satisfied.
\todo{Détailler la contraite 21 ou l'enlever} 
\begin{lstlisting}
satisfiable(F, T) :-  goal(F), instate(F, T). %\label{c2}
:- X={ satisfaible(F, T+1) : goal(F) }, getNbreGoals(Y), X=Y, time(T). %\label{c2-a}
\end{lstlisting}

>>>>

\subsection{Reachability verification}
In this section, we focus on the reachability of set of processes which corresponds to the reachability question (see definition \ref{def:reachability}):
“Is it possible, starting from a given initial state, to play a number of actions so that a set of given processes are active in the resulting state?”
We now want to adapt the code of the dynamics computation of previous section in order to resolve this reachability problem.
For this, we first use a predicate to list the processes we want to reach, called \texttt{goal}, and we add as many rules of the following form as there are objective processes:
\begin{lstlisting}
goal(activeProcess("a",1)). %\label{c1}
\end{lstlisting}
Then, the literal \texttt{satisfiable(F, T)} 
checks if a given process \texttt{F} of the goal
is contained in the state of step \texttt{T},
as defined in the rule of line \ref{c2}.
Else the answer will be eliminated by the constraint \ref{c2-a} which verifies if all processes of the goal are satisfied.
\todo{Détailler la contraite 21 ou l'enlever} 
\begin{lstlisting}
satisfiable(F, T) :-  goal(F), instate(F, T). %\label{c2}
:- X={ satisfaible(F, T+1) : goal(F) }, getNbreGoals(Y), X=Y, time(T). %\label{c2-a}
\end{lstlisting}
